---
title: "181058 DSLab Final"
output:
  html_document: default
  pdf_document: default
---

> Task 01

>> Reading the dataset

importing dataset into R

```{r}
df = read.csv('regression_Data.csv')
head(df)
```

 checking for its structure to find nore information about the given dataset

```{r}
str(df)
```
From the output, we can see that we have 9568 observations/ rows for 5 variables/column. We can also see that all are of types _num_

>> Cleaning the data

For cleaning the data, we need to

1. Finding outliers 
2. Finding null values 
3. Finding categorical values

>> Finding outliers

Best method is to use boxplot, values outside the plot are outliers
```{r fig.width=8, fig.height=8}
boxplot(df)
```
From the figure, we don't seem to find any outlier, trying again, mathematically.
1. We could find  mean and standard deviation, from where we may find Q1 and Q3, Q2 is mean
Expression for outliers will be 
_Quartile = 1.5(Q3-Q1)_
Values below mean and above max would be classified as outliers which should be removed.
```{r}
quartiles <- quantile(df$AT, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(df$AT)
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
#finding outlier
df$AT[which(df$AT < Lower | df$AT > Upper)]


quartiles <- quantile(df$V, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(df$V)
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
#finding outlier
df$V[which(df$V < Lower | df$V > Upper)]

quartiles <- quantile(df$AP, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(df$AP)
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
#finding outlier
df$AP[which(df$AP < Lower | df$AP > Upper)]

quartiles <- quantile(df$RH, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(df$RH)
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR
#finding outlier
df$RH[which(df$RH < Lower | df$RH > Upper)]


quartiles <- quantile(df$PE, probs=c(.25, .75), na.rm = FALSE)
IQR <- IQR(df$PE)
Lower <- quartiles[1] - 1.5*IQR
Upper <- quartiles[2] + 1.5*IQR 
#finding outlier
df$PE[which(df$PE < Lower | df$PE > Upper)]

```

So, we have outliers in AP and RH
Removing them.
```{r}
without_outlier <- subset(df, df$AP > Lower & df$AP < Upper)
without_outlier <- subset(df, df$RH > Lower & df$RH < Upper)

dim(without_outlier)

outliers <- boxplot(df, plot = FALSE)$out
data_no_outlier <- df[-which(df %in% outliers)] 
dim(data_no_outlier)
length(data_no_outlier)

```

> Task 02 Linear Regression 

>> Plotting dependant and independant variables

Finding corelation between all variables
```{r}
#installing corrplot
library(corrplot)
corr <- sapply(df, is.numeric)
corr

cor.data <- cor(df[, corr])
cor.data

corrplot(cor.data, method='color')

```

>> Splitting the data
```{r}
#Splitting the data
library(caTools)
set.seed(2)
split <- sample.split(df, SplitRatio = 0.8)
train <- subset(df, split == 'TRUE')
test <- subset(df, split == 'FALSE')

dim(train)
dim(test)
```
>> Linear Regression, plotting and summary
```{r}
#making a model
model <- lm(PE ~., data = train)
summary(model)
pred <- predict(model, test)
head(pred)

res <-residuals(model)
res <- as.data.frame(res)
head(res)
```
```{r, message=FALSE, warning=FALSE}
#Compairing with actual values
results <- cbind(pred, df$PE)
colnames(results) <- c('Predicted', 'Real')
results <- as.data.frame(results)
head(results)

plot(df$PE, type= 'l')

par(mfrow=c(2,2))
plot(model)
mtext("Linear Model",                  
      side = 3,
      line = - 1,
      outer = TRUE)
summary(head((model)))

```

>> Polynomial regression with degree 2 and 3 and summary 

```{r, message=FALSE, warning=FALSE}
DegreeModel2 <- lm(df$PE ~ poly(df$AP, 2), data = train)
par(mfrow=c(2,2))
plot(DegreeModel2)
mtext("Degree Model 2",                  
      side = 3,
      line = - 1,
      outer = TRUE)

summary(head((DegreeModel2)))

DegreeModel3 <- lm(df$PE ~ poly(df$AP, 2), data = train)
par(mfrow=c(2,2))
plot(DegreeModel3)
mtext("Degree Model 3",                  
      side = 3,
      line = - 1,
      outer = TRUE)

summary(head((DegreeModel3)))

```

>> Predicting the result using trained data

```{r, message=FALSE, warning=FALSE}
predicted_result <- predict(model, test)

res <- residuals(model)
res <- as.data.frame(res)

final_res <- cbind(predicted_result, df$PE)

colnames(final_res) <- c("Predicted","real")

final_res <- as.data.frame(final_res)

#plotting them
plot(df$PE, type = 'l', lty = 1.8, col = 'red')
lines(predicted_result, col='green')

#checking accuracy
rmse <- sqrt(mean(predicted_result - df$PE)^2)
paste("Accuracy is ", rmse)
```
