---
title: "181058_DSLab_MID"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

>Task 01

Create a Vector and save 6 values 1 ,2, 3,4,5,6 your id digits, Assign names to each instance and display length and type of vector.

```{r}
ID <- c(1,8,1,0,5,8)
names(ID) <- c("one","eight","one", "zero", "five","eight")
# Can also do it like 
# id <- c("One"=1,"Eight"=8,"One"=1, "Zero"=0, "Five"=5,"Eight"=8)
print(ID)
paste("LENGTH -> ", length(ID)) #alternative for concatenating
paste("Type -> ", typeof(ID)) 
```

>Task 02

Create a matrix "YourName" and store 9 values row wise with 3 rows and name the row and column names. Multiply Complete Matrix with 100 and Display the output.

```{r}
HamzaAli <- matrix(1:9, nrow = 3, byrow=TRUE)
rownames(HamzaAli) <- c("R1","R2","R3")
colnames(HamzaAli) <- c("C1","C2","C3")
new_mat <- HamzaAli * 100
print(HamzaAli)
print(new_mat)
```

>Task 03

Using Provided Dataset

```{r}
df <- read.csv("DataSet_MidExam_C.csv")
print("STRUCTURE")
print(str(df))
print("SUMMARY")
print(summary(df))
```
#### Str vs Summary
According to the Output of Str and Summary, 

1. Str provides the following:
  +   The total number of rows and columns, i.e. 1216 rows(objects) and 7 columns(variables)
  +   "$" gives description of each column(their factor(unique values) and int values itselfs)
  
2. As for summary, 
  +   It gives summary of head, 6 rows and others
  +   Each column would gives total number of distinct rows, for example, Barbados has 44 rows in Country...territory
  +   For Integer, it would gives out Min, max, media, median, number of NA's and basic math functions
  +   In short, it would give the summary of the dataset to give bried description of the provided datasets
  
#### Finding the Categorical and Numerical attributes 
  *   Any int value gives numerical attributes
  *   Any factor of limited values (M/F, Confirmed/Suspected) gives Categorical attribues
Now, looking at summary, we can draw following conclusion
  1.  **Country...territory**, **Measure.Names**, **Measure.Values**, **Suspected.cases** has distinct values(factor), so it's categorical attributes. **Year.of.Data** gives 2 values (2015,2016), so it can be also categorical 
  2.  **Report.Epi.Week**, **Year.of.Date**, **Laboratory.Confirmed.cases**, **Suspected.cases** are all numerical attributes
  
```{r}
df <- read.csv("DataSet_MidExam_C.csv")
plot(main = "Categorical PLOT", df$Country...territory, df$Measure.Names)
plot(main = "Numerical Plot", df$Report.Epi.Week, df$Year.of.Date)
plot(main = "2X", (df$Report.Epi.Week)*2, (df$Year.of.Date)*2, xlab = "X", ylab="Y")
hist(df$Report.Epi.Week,main = "Histogram of Report EPI WEEK", col="green", breaks = 10)
```



>Task 04

#### Top & Last5 entries of Provided Dataset
```{r}
print("Top 5 values")
print(head(df, n = 5))
print("Last 5 values")
print(tail(df, n = 5))
print("Summary")
print(summary(df))
print("5 Line Summary")
fivenum(df$Report.Epi.Week)
fivenum(df$Year.of.Date)
fivenum(df$Laboratory.confirmed.cases)
```
#### Difference Between Summary and 5Num summary

1. As described by its name, summary gives detailed view of the datasets, even __factors__ and __string__
2. Factor only works with __numerical attributes__ as it gives provides 
* Minimum
* First Quartile
* Median
* Third Quartile
* Maximum


#### BoxPlot and Outliers

```{r}
print("BoxPlot")
boxplot(df$Report.Epi.Week)
```
 According to the Boxplot, 
 * Lowest line gives min values = 1
 * 2nd lower will gives 1st quartile = 5
 * bolded line is the median of column = 43 
 * Above that gives third quartilse = 48 
 * Top values give Maximun values = 52
 All the values below minimum and above maximum are referred as outliers 
 (1 > Outliers > 52)
 
 >Task 05

#### Encode Categorical to Numerical
```{r}
df <- read.csv("DataSet_MidExam_C.csv")
df$Country...territory <- factor(df$Country...territory, labels= 1:31)
df$Measure.Names <- factor(df$Measure.Names, labels = c(1,2))
df$Measure.Values <- factor(df$Measure.Values, labels = 1:113)
df$Suspected.cases <- factor(df$Suspected.cases, labels = 1:88)
```

#### Missing values with mean
According to summary, we have NA at Report.Epi.Week
```{r}
df$Report.Epi.Week <- ifelse(is.na(df$Report.Epi.Week),
                             ave(df$Report.Epi.Week, FUN = function(x) mean(x, na.rm = TRUE)), df$Report.Epi.Week)
#print(df$Report.Epi.Week)
#for conforming
#is.na(df$Report.Epi.Week)
```
#### Split to 80/20
```{r}
library(caTools)
set.seed(123)
split = sample.split(df$Measure.Values, SplitRatio = 0.8)
trainingset <- subset(df, split == TRUE)
testset <- subset(df, split == FALSE)
paste("Training Set", dim(trainingset)) 
paste("Test set", dim(testset))

```